---
title: "A/B Testing"
author: "Jake Hofman"
date: "April 17, 2015"
output:
  html_document:
    toc: true
    toc_depth: 2
---


```{r}
library(ggplot2)
library(reshape)

theme_set(theme_bw())

set.seed(42)
```


# Estimating a proportion
## Point estimate and sampling distribution
Repeatedly flip a biased coin 100 times and estimate its bias.
Taken from Yakir 11.2.3.
```{r}
p <- 0.3
n <- 100
P.hat <- rep(0,10^5)
for(i in 1:10^5) {
  X <- rbinom(n,1,p)
  P.hat[i] <- mean(X)
}

# plot the sampling distribution
qplot(x=P.hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=p) +
  geom_vline(xintercept=mean(P.hat), linetype=2, color="red")
```

## Confidence interval
```{r}
# compute upper and lower confidence intervals
LCL <- P.hat - 1.96*sqrt(P.hat*(1-P.hat)/n)
UCL <- P.hat + 1.96*sqrt(P.hat*(1-P.hat)/n)

# check how often the true proportion is contained in the estimated confidence interval
mean(p >= LCL & p <= UCL)

# plot 100 confidence intervals and the true value
plot_data <- data.frame(P.hat, LCL, UCL)[1:100,]
plot_data <- transform(plot_data, contains_p=(p >= LCL & p <= UCL))
ggplot(data=plot_data, aes(x=1:nrow(plot_data), y=P.hat, color=contains_p)) +
  geom_pointrange(aes(ymin=LCL, ymax=UCL)) +
  geom_hline(yintercept=p, linetype=2) +
  xlab('') +
  scale_color_manual(values=c("red","darkgreen")) +
  theme(legend.position="none")
```


## Hypothesis testing
```{r}
# construct a null distribution: what would happen if the coin were fair?
p <- 0.5
n <- 100
P0.hat <- rep(0,10^5)
for(i in 1:10^5) {
  X <- rbinom(n,1,p)
  P0.hat[i] <- mean(X)
}

# plot the null distribution and see where the observed estimate lies in it
qplot(x=P0.hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=P.hat[1], linetype=2, color="red")

# compare this to one of our experiments with the biased coin above
# how likely is it that we would see an estimate this extreme if the coin really were fair?
num.as.extreme <- sum(P0.hat <= P.hat[1])
p.value <- num.as.extreme / length(P.hat)
```
Only `r num.as.extreme` out of `r length(P.hat)` flips of a fair coin with p=0.5 would result in an estimate of p_hat=`r P.hat[1]` or smaller, corresponding to a p-value of `r p.value`.

# Comparing two proportions
## Point estimates and sampling distributions
Repeatedly flip two coins, each 500 times and estimate their bias.
```{r}
pa <- 0.12
pb <- 0.08
n <- 500
PA.hat <- rep(0,10^5)
PB.hat <- rep(0,10^5)
for(i in 1:10^5) {
  XA <- rbinom(n,1,pa)
  PA.hat[i] <- mean(XA)
  XB <- rbinom(n,1,pb)
  PB.hat[i] <- mean(XB)
}

# plot the sampling distribution of the difference
qplot(x=PA.hat-PB.hat, geom="histogram", binwidth=0.002) +
  geom_vline(xintercept=pa-pb) +
  geom_vline(xintercept=mean(PA.hat-PB.hat), linetype=2, color="red")

# wrangle the results into one data frame
PA.hat <- data.frame(PA.hat)
PA.hat$split <- 'A'
PA.hat$trial <- 1:nrow(PA.hat)
names(PA.hat) <- c('P.hat', 'split', 'trial')
PB.hat <- data.frame(PB.hat)
PB.hat$split <- 'B'
PB.hat$trial <- 1:nrow(PB.hat)
names(PB.hat) <- c('P.hat', 'split', 'trial')
plot_data <- rbind(PA.hat, PB.hat)

# plot the sampling distributions for each split
qplot(data=plot_data, x=P.hat, fill=split, geom="histogram", position="identity", binwidth=0.002)
```

## Confidence intervals
```{r}
# plot 100 confidence intervals by split
plot_data <- transform(plot_data, 
                       LCL = P.hat - 1.96*sqrt(P.hat*(1-P.hat)/n),
                       UCL = P.hat + 1.96*sqrt(P.hat*(1-P.hat)/n))
plot_data <- subset(plot_data, trial <= 100)
ggplot(data=plot_data, aes(x=trial, y=P.hat, linetype=split, position="dodge")) +
  geom_pointrange(aes(ymin=LCL, ymax=UCL)) +
  xlab('') +
  theme(legend.title=element_blank())
```

## Hypothesis testing
```{r}
# construct a null distribution: what would happen if both coins had the same bias (e.g., A and B are the same)?
p0a <- 0.08
p0b <- 0.08
n <- 500
dP0.hat <- rep(0,10^5)
for(i in 1:10^5) {
  XA <- rbinom(n,1,p0a)
  XB <- rbinom(n,1,p0b)
  dP0.hat[i] <- mean(XA-XB)
}

# run one experiment where there is an underlying difference
pa <- 0.12
pb <- 0.08
XA <- rbinom(n,1,pa)
XB <- rbinom(n,1,pb)
dP.hat <- mean(XA - XB)

# plot the null distribution and see where the observed estimate lies in it
qplot(x=dP0.hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=dP.hat, linetype=2, color="red")

# compare this to one of our experiments with the biased coin above
# how likely is it that we would see an estimate this extreme if the coin really were fair?
num.as.extreme <- sum(dP0.hat >= dP.hat)
p.value <- num.as.extreme / length(dP0.hat)
```
Only `r num.as.extreme` out of `r length(P.hat)` flips of two identical coins with p=0.08 would result in an estimate of dp_hat=`r dP.hat` or larger, corresponding to a p-value of `r p.value`.
